SunPy - Python for Solar Physics

### 2014-01-29 2-2-Lightcurve.tex
my changes

Time series data and their analyses are a fundamental aspect of solar
physics for which many data sources are available.
SunPy provides a \texttt{LightCurve} class
with a convenient and consistent interface for handling solar time-series
data.  The main engine behind the \texttt{LightCurve} class is
the \href{http://pandas.pydata.org}{\texttt{pandas}} data-analysis library, and 
\texttt{LightCurve}'s \texttt{data} attribute is a \texttt{pandas.DataFrame} 
class.
The \texttt{pandas} library contains a large amount
of functionality for manipulating and analysing time-series data,
making it an ideal basis for \texttt{LightCurve}.  \texttt{LightCurve}
assumes that the input data are time-ordered list(s) of numbers, and each
list becomes a column in the \texttt{pandas} DataFrame class.

Currently, the \texttt{LightCurve} class is compatible with the
following data sources: the GOES X-ray Sensor (XRS), PROBA2/LYRA, and
the SDO EUV Variability Experiment (EVE; only the level ``OCS'' and
% Do we need to explain what OCS means?
% RJH: I'd argue that everything other than (EVE) can be cut, but I'm not 
% confident enough to make that cut.
average CSV files -- see \url{http://lasp.colorado.edu/home/eve/data/}
for more detail).  For each of these instruments, a subclass of the
\texttt{LightCurve} object is initialised
(\textit{e.g.}, \texttt{GOESLightCurve}) which inherits from
\texttt{LightCurve} but allows instrument-specific functionality to be
included.  Future developments will introduce support for additional
instruments and data products, as well as implementing a factory interface 
similar to that of \texttt{Map}.  Since there is no established standard
as to how time-series data should be stored and distributed, each SunPy 
\texttt{LightCurve} object sub-class provides the ability to download its corresponding 
specific data format in its constructor and parse that file type.

@@

\begin{listing}[H]
\begin{minted}[bgcolor=bg]{pycon}
>>> from sunpy import lightcurve
>>> from sunpy.time import TimeRange
>>> goes = lightcurve.GOESLightCurve.create('2011-06-07 06:00',
...                                         '2011-06-07 08:00')
>>> goes.peek()
%removed the resample part of the example - really we need to fix truncate!
\end{minted}
\begin{center}
\includegraphics[width=10cm]{goes_lightcurve.pdf}
\end{center}
\caption{Example retrieval of a GOES lightcurve for the time interval 
06:00--08:00 UT on 2011 June 7 using a time range, and the output of the 
\texttt{peek()} method.}
\label{code:goes_lc}
\end{listing}
%schriste - note to self, fix this example, add resampled points on top of plot?

>>>>

Time series data and their analyses are a fundamental aspect of solar
physics for which many data sources are available.
SunPy provides a \texttt{LightCurve} class
with a convenient and consistent interface for handling solar time-series
data.  The main engine behind the \texttt{LightCurve} class is
the {\texttt{pandas}} data-analysis library (\url{http://pandas.pydata.org}), and 
\texttt{LightCurve}'s \texttt{data} attribute is a \texttt{pandas.DataFrame} 
class.
The \texttt{pandas} library contains a large amount
of functionality for manipulating and analysing time-series data,
making it an ideal basis for \texttt{LightCurve}.  \texttt{LightCurve}
assumes that the input data are time-ordered list(s) of numbers, and each
list becomes a column in the \texttt{pandas} DataFrame class.

Currently, the \texttt{LightCurve} class is compatible with the
following data sources: the GOES X-ray Sensor (XRS), PROBA2/LYRA, and
the SDO EUV Variability Experiment (EVE; only the level ``OCS'' and
% Do we need to explain what OCS means?
% RJH: I'd argue that everything other than (EVE) can be cut, but I'm not 
% confident enough to make that cut.
average CSV files -- see \url{http://lasp.colorado.edu/home/eve/data/}
for more detail).  For each of these instruments, a subclass of the
\texttt{LightCurve} object is initialised
(e.g., \texttt{GOESLightCurve}) which inherits from
\texttt{LightCurve} but allows instrument-specific functionality to be
included.  Future developments will introduce support for additional
instruments and data products, as well as implementing a factory interface 
similar to that of \texttt{Map}.  Since there is no established standard
as to how time-series data should be stored and distributed, each SunPy 
\texttt{LightCurve} object sub-class provides the ability to download its corresponding 
specific data format in its constructor and parse that file type.

@@

\begin{listing}[H]
\begin{minted}[bgcolor=bg]{pycon}
>>> from sunpy import lightcurve
>>> from sunpy.time import TimeRange
>>> goes = lightcurve.GOESLightCurve.create('2011-06-07 06:00',
...                                         '2011-06-07 08:00')
>>> goes.peek()
%removed the resample part of the example - really we need to fix truncate!
% schriste - i think it would be better to show off some of the powerful pandas methods in this section
% as we are not likely to fix truncate and i think it should be refactored anyway
\end{minted}
\begin{center}
\includegraphics[width=10cm]{goes_lightcurve.pdf}
\end{center}
\caption{Example retrieval of a GOES lightcurve for the time interval 
06:00--08:00 UT on 2011 June 7 using a time range, and the output of the 
\texttt{peek()} method.}
\label{code:goes_lc}
\end{listing}
%schriste - note to self, fix this example, add resampled points on top of plot?

### 2014-01-29 2-3-Spectra.tex
my changes

\subsection{Spectra}\label{sec:spectra}
%schriste - this section needs major work
%ji - there is no example of a spectrum
%dps - I need to think in one...

@@

The \texttt{Spectrogram} class supports radio spectrograms from the 
\href{http://www.e-callisto.org/}{e-Callisto}
solar radio spectrometer network and STEREO/SWAVES spectrograms.
As with other SunPy data types, the \texttt{Spectrogram} class has been
built so that each instrument initialises using a subclass containing the instrument-specific 
functionalities.
%The common functionality provided by the base \texttt{Spectrogram} class includes
%reading data,
%joining different time ranges and frequencies,
%performing frequency-dependent background subtraction,
%and convenient visualization and sampling of the data.

Listing \ref{code:spectra} shows how the \texttt{CallistoSpectrogram}
object retrieves spectrogram data in the time range specified taken at
the observatory of interest.  When the data is requested using the
\texttt{from\_range} function, the object merges all the downloaded
files into a single spectrogram, across time and frequency.
In the example shown, data is provided in two frequency ranges, 
20--90\,MHz and 55--355\,MHz.  Since the data is not evenly spaced in
the frequency range, the \texttt{Spectrogram} object linearises the
frequency axis for a better analysis.  The example also demonstrates
the implemented background subtraction method.
% RJH: Which subtraction method?  If the specifics are irrelevant, cut it.
%ji to DPS - what is the linearisation that is happening here?

>>>>

\subsection{Spectra}\label{sec:spectra}
%schriste - this section needs major work
%ji - there is no example of a spectrum
%dps - I need to think in one...
%schriste - can't we just take a slice off of the spectram to make a spectrum?

@@

As with other SunPy data types, the \texttt{Spectrogram} class has been
built so that each instrument initialises using a subclass containing the instrument-specific 
functionalities. The common functionality provided by the base \texttt{Spectrogram} class includes
joining different time ranges and frequencies, performing frequency-dependent background subtraction,
and convenient visualization and sampling of the data.
Currently, the \texttt{Spectrogram} class supports radio spectrograms from the 
\href{http://www.e-callisto.org/}{e-Callisto}
solar radio spectrometer network and STEREO/SWAVES spectrograms.

Listing \ref{code:spectra} shows how the \texttt{CallistoSpectrogram}
object retrieves spectrogram data in the time range specified taken at
the observatory of interest.  When the data is requested using the
\texttt{from\_range()} function, the object merges all the downloaded
files into a single spectrogram, across time and frequency.
In the example shown, data is provided in two frequency ranges, 
20--90\,MHz and 55--355\,MHz.  Since the data is not evenly spaced in
the frequency range, the \texttt{Spectrogram} object linearises the
frequency axis for a better analysis.  The example also demonstrates
the implemented background subtraction method.
% RJH: Which subtraction method?  If the specifics are irrelevant, cut it.
%ji to DPS - what is the linearisation that is happening here?

### 2014-01-29 2-Data_Types.tex
my changes

Currently, SunPy provides data structures that are specifically designed for the
three primary varieties of solar physics data: images, time series, and
spectra. SunPy provides core data types for these using Python classes:
\texttt{Map} (2D spatial data), \texttt{LightCurve} (1D temporal series)
and \texttt{Spectrum} and \texttt{Spectrogram} (1D and 2D spectra).

These classes allow access to the original data
along with associated metadata and provide appropriate convenience functions to
enable data analysis and visualisation. For each of these classes, the data is
stored in the \texttt{data} attribute while the metadata is stored 
in the \texttt{meta} attribute. (Curently, \texttt{LightCurve} does 
not support the \texttt{meta} attribute.) It is possible to instantiate the
data types from various
different sources: e.g., files, URLs, arrays or time ranges.  In order 
to provide instrument-specific specialisation, all the core SunPy data types 
support subclassing; e.g., \texttt{Map} has an \texttt{AIAMap} 
sub-type for data from the SDO/AIA instrument. 
Data visualisation is provided by two functions: \texttt{peek()}, for quick 
plotting, and \texttt{plot()}, for plotting with more fine-grained control that 
integrates with \texttt{matplotlib}.

@@

This design philosophy makes the behaviour of SunPy's visualisation 
routines intuitive to those who already understand the \texttt{matplotlib}
interface, as well as allowing the use of the standard 
\texttt{matplotlib} commands to manipulate the plot parameters, such as the 
title.

>>>>

SunPy provides data strucutures that are specifically designed for the
three primary varieties of solar physics data: images, time series, and
spectra. These core data types are Python classes:
\texttt{Map} (2D spatial data), \texttt{LightCurve} (1D temporal series)
and \texttt{Spectrum} and \texttt{Spectrogram} (1D and 2D spectra).

These classes allow access to the data
along with associated metadata and provide appropriate convenience functions to
enable analysis and visualisation. For each of these classes, the data is
stored in the \texttt{data} attribute while the metadata is stored 
in the \texttt{meta} attribute. (Curently, \texttt{LightCurve} does 
not support the \texttt{meta} attribute.) 
% schriste - can't we just quickly fix this? should be easy enough.
It is possible to instantiate the
data types from various
different sources: e.g., files, URLs, arrays.  
%schriste - i removed time ranges as that just looks up a url.
In order to provide instrument-specific specialisation, all the core SunPy classes 
support subclassing; e.g., \texttt{Map} has an \texttt{AIAMap} 
sub-type for data from the SDO/AIA instrument. 
Data visualisation is provided by two functions: \texttt{peek()}, for quick 
plotting, and \texttt{plot()}, for plotting with more fine-grained control that 
integrates with \texttt{matplotlib}.

@@

This design philosophy makes the behaviour of SunPy's visualisation 
routines intuitive to those who already understand the \texttt{matplotlib}
interface, as well as allowing the use of the standard 
\texttt{matplotlib} commands to manipulate the plot parameters (e.g. title, axes).

### 2014-01-29 3-2-HEK.tex
closes issue #75
my changes

The Sun is an active star and exhibits a wide range of transient phenomena 
(e.g., flares and radio bursts) on many different time-scales, length-scales and 
wavelengths. Observations and metadata concerning these phenomena are collected 
in the Heliophysics Event Knowledgebase (HEK) [REF].  Entries are generated both by 
automated algorithms and human observers.  Some of the information in the HEK 
reproduces feature and event data from elsewhere (for example, the GOES flare catalogue),
and some is generated by the Solar Dynamics Observatory Feature Finding Team 
\citep{martens2012}.  The advantage of the HEK is it 
provides an homogeneous and well-described interface to a large amount of 
feature and event information of interest to the solar physics community.

SunPy accesses this information through the \texttt{hek} module, which was also
developed with support from ESA's SOCIS 2011.  The \texttt{hek} module makes 
use of the 
\href{http://vso.stanford.edu/hekwiki/ApplicationProgrammingInterface?action=print}{HEK
 API}.
The current list of events maintained by the HEK and their properties can be 
found at \url{http://www.lmsal.com/hek/VOEvent_Spec.html}.

>>>>

The Sun is an active star and exhibits a wide range of transient phenomena 
(e.g., flares and radio bursts) on many different time-scales, length-scales and 
wavelengths. Observations and metadata concerning these phenomena are collected 
in the Heliophysics Event Knowledgebase (HEK) \citep{hek2012}.  Entries are generated both by 
automated algorithms and human observers.  Some of the information in the HEK 
reproduces feature and event data from elsewhere (for example, the GOES flare catalogue),
and some is generated by the Solar Dynamics Observatory Feature Finding Team 
\citep{martens2012}.  The advantage of the HEK is it 
provides an homogeneous and well-described interface to a large amount of 
feature and event information of interest to the solar physics community.

SunPy accesses this information through the \texttt{hek} module, which was
developed with support from ESA's SOCIS 2011.  The \texttt{hek} module makes 
use of the 
\href{http://vso.stanford.edu/hekwiki/ApplicationProgrammingInterface?action=print}{HEK
 API}.
The current list of events maintained by the HEK and their properties can be 
found at \url{http://www.lmsal.com/hek/VOEvent_Spec.html}.
